---
title: "100 nm Al on Fe 20kV base case 1000 ch"
author: "John Minter"
date: 'Started 2018-08-07, Last modified: 2018-08-13'
output:
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    css: ../../../../inc/jm.css
    number_sections: yes
    toc: yes
abstract: |
  The objective was to analyze the base case for 100 nm of Al on Fe at
  15 kV using the base case from Llovet and Salvat 2018 table 3 case 1
  using the Al K-L3 ($K_{\alpha\ 1}$) transition. In this first
  (baseline) simulation I only collected 1000 channels of data.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
  comment = NA,
  fig.align = "centre",
  fig.height = 4,
  message = FALSE,
  warning = FALSE,
  error = FALSE)
```

# Set up for the analysis

**Note**: This simulation used the same `MSIMPA` parameters for all layers
and a `REFLIN` parameter (`5E-1`) to finish.

```
MFNAME Al.mat                         [Material file, up to 20 chars]
MSIMPA 1.0E+3 1.0E+3 1E+3 0 0 0 -1E+2       [EABS(1:3),C1,C2,WCC,WCR]
MFNAME Fe.mat                         [Material file, up to 20 chars]
MSIMPA 1.0E+3 1.0E+3 1E+3 0.2 0.2 1 1       [EABS(1:3),C1,C2,WCC,WCR]
```

First load the libraries we need...

```{r loadPackages}
library(rpenepma)
library(dplyr)
library(pander)
panderOptions('table.split.table', Inf)
library(here)
library(ggplot2)
```

Next, set the paths and constants that we need...

```{r setPaths}
moved_data  <- TRUE  # set to TRUE when sim is done and .rda with data is moved...
e0          <- 15 # kV
int_lo_lim  <- 1.0e-1
int_hi_lim  <- 1.0e4
reflin_crit <- 5e-1
sim_dir <- here("penepma/penepma16/Al-100nm-on-Fe/Al_100_Fe_15kV_1000_ch")

sim_nam     <- "Al-100nm-on-Fe-base-1000-channels"
res_fi      <- "./penepma-res.dat"
int_fi      <- "./pe-intens-01.dat"
spc_fi      <- "./pe-spect-01.dat"
out_ti      <- sprintf("%s-%gkV", sim_nam, e0)
msa_fi      <- sprintf("./%s-%gkV.msa", sim_nam, e0)
```


# Get the current simulation time

```{r calcSimTime}
if (moved_data==FALSE){
  # When we have not moved the simulation, measure the simulation time.
  # Note copying the files to another directory seems to change the time
  # Test a reload.
  delta_t <- penepma_measure_simulation_time(sim_dir)
  # print(delta_t)
  save(delta_t, file="./delta_t.rda")
  rm(delta_t)
  load("./delta_t.rda")
  msg <- "Analyzed from simulation directory."
} else {
  # When we have moved the simulation, reload the delta_t.rda
  msg <- "Analyzed from data stored in delta_t.rda."
  load("./delta_t.rda")	
}
cat(c(msg, "\nCurrent simulation time:",
      round(as.numeric(delta_t, units="hours"), 3), "hrs"))

```

# Retrieve the number of showers

We want retrieve the number of showers (trajectories) that our `penepma16`
simulation calculated.

```{r calcNumShowers}
options(scipen = -6) # force printing in exponential format
num_showers <- penepma_get_number_of_showers(res_fi)
options(scipen = 3) # reset to default
cat("Number of trajectories: ", num_showers)
```

# Compute the total maximum intensity for each element

## Retrieve the data

First retrieve the intensity data as a `tibble` and print a preview.

```{r LoadMaxIntensity}
tib <- penepma_get_intensities(int_fi)
print(tib)
```

What features did we measure?

```{r listFeatures}
print(names(tib))
```

How many transitions did we measure?

```{r calcNumTrans}
print(nrow(tib))
```

Measure our test transition ("Al K-L3")

```{r measureZnTrans}
val <- penepma_get_total_intensity_z_transition(tib, 13, "K", "L3")
print(val)
```



## Compute the maximum total intensity for each element

1. Compute the maximum total intensity for each element.
2. Stack each row of data together into a tibble.
3. Prepend the sample ID to the data.
4. Print a well-formatted a table using `pander`.
5. Save the final tibble so it can be r-loaded later

```{r calcNaxTotIntensity}
# 1
Al_t   <- penepma_get_max_total_intensity_z(tib, 13)
Fe_t   <- penepma_get_max_total_intensity_z(tib, 26)
# 2
tot_int <- dplyr::bind_rows(Al_t, Fe_t)
#3 
Al_100_Fe_15kV_1000_ch <- prepend_sample_id_max_int_tib(tot_int, out_ti)
# 4
pander(Al_100_Fe_15kV_1000_ch)
# 5
save(Al_100_Fe_15kV_1000_ch,
     file="Al_100_Fe_15kV_1000_ch.rda")
# delete and reload be sure it worked...
rm(Al_100_Fe_15kV)
load("./Al_100_Fe_15kV_1000_ch.rda")
```

# Process the spectra data

## Load the raw data

```{r loadSpectraData}
tib <- penepma_read_raw_data(spc_fi)
rownames(tib) <- c()
```

## Summarize the spectrum data

How many rows of data are there?

```{r cat_num_rows}
cat(c("Number of measurements: ", nrow(tib)))
```

Generate a summary:

```{r summarizeSpectrum}
pander(summary(tib))
```

## Plot the spectrum.

Start with a log-intensity plot.


```{r plotLogIntensSpectrum, fig.width=8, warning=FALSE}
m_t <- max(pretty(tib$mu))

plt <- ggplot(tib, aes(x = keV, y = pd.mu)) +
       geom_line() + 
       scale_x_continuous(breaks = seq(from = 0, to = e0-5, by = 1),
                          limits = c(0,e0-5)) +
       scale_y_log10() +
       xlab(label="X-ray energy [keV]") +
       ylab(label="log(probability density)") +
       # (1/(eV*sr*electron)") +
       ggtitle(out_ti) +
       theme(axis.text=element_text(size=12),
             axis.title=element_text(size=12),
             # center the title
             plot.title = element_text(hjust = 0.5))
       
print(plt)
```
And then on a linear intensity scale...

```{r plotLinIntensSpectrum, fig.width=8, warning=FALSE}

plt <- ggplot(tib, aes(x = keV, y = pd.mu)) +
       geom_line() + 
       scale_x_continuous(breaks = seq(from = 0, to = e0-5, by = 1),
                          limits = c(0,e0-5)) +
       scale_y_continuous() +
       xlab(label="X-ray energy [keV]") +
       ylab(label="Probability density") +
       # (1/(eV*sr*electron)") +
       ggtitle(out_ti) +
       theme(axis.text=element_text(size=12),
             axis.title=element_text(size=12),
             # center the title
             plot.title = element_text(hjust = 0.5))
       
print(plt)
```

## Write a spectrum file in MSA format

```{r writeMsa}
penepma_to_msa(spc_fi, msa_fi,e0, out_ti)
```


