---
title: "Analyze pypenelope simulations of C on Fe with different Fe density values"
author: "John Minter"
date: 'Started 2018-09-17, Last modified: 2018-09-21'
output:
  html_document:
    css: ./jm-gray-vignette.css
    number_sections: yes
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
abstract: |
  The objective was to test an idea suggested by Aurien Moi. Aurien reported
  that he had run penepma simulations on Fe with densities set to 0.1 and 10.0
  g/cm3 and that the reported intensities were within experimental error. This
  was in the context of my question of whether the GMRFILM program not using the
  substrate density was a problem. I wanted to test this on a thin film case 
  with a low Z coating, so I simulated C on Fe at 20 kV.
  
  More to come...
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
  comment = NA,
  fig.align = "centre",
  fig.height = 4,
  message = FALSE,
  warning = FALSE,
  error = FALSE)
```

# Set up for the analysis

This report was generated from a Rmarkdown file that is supplied. The R code
chunks contain the code to reproduce the analysis from the data files that are
supplied. To make this report more readable, only the key output is rendered in
the HTML report. The very first chunk (named `setup`) sets some general options.
The second code chunk (named `loadPackages`) loads the R packages that ontain
the functions to process the penepma data files to extract the intensities and
permit us to compute the K-ratios. The third code chunk (named `setPaths`) 
sets the electron beam energy to 20 kV and generates the file paths for each of
the penepma intensity files. These were the `pe-inten-01.dat` file for each 
specimen with specimen identifier string prepended.

```{r loadPackages}
library(here)
library(rpenepma)
library(dplyr)
library(pander)
panderOptions('table.split.table', Inf)
library(ggplot2)
```


```{r setPaths}
e0          <- 20 # kV
sim_dir     <- here()
c_std   <- sprintf("%s/c-std-pe-inten-01.dat", sim_dir)
fe_std_p1   <- sprintf("%s/fe-den-p1-pe-inten-01.dat", sim_dir)
fe_std_10   <- sprintf("%s/fe-den-10-pe-inten-01.dat", sim_dir)
c20nm_fe_p1   <- sprintf("%s/c-20-nm-on-fe-den-p1-pe-inten-01.dat", sim_dir)
c20nm_fe_10   <- sprintf("%s/c-20-nm-on-fe-den-10-pe-inten-01.dat", sim_dir)
c200nm_fe_p1   <- sprintf("%s/c-200-nm-on-fe-den-p1-pe-inten-01.dat", sim_dir)
c200nm_fe_10   <- sprintf("%s/c-200-nm-on-fe-den-10-pe-inten-01.dat", sim_dir)
print(c_std)
```

# Compute the intensities need to compute K-ratios

Next, in the code chunk `createIntensTibbles`, we retrieve the intensity
for the C and Fe transitions of interest
(C Ka1, FeLa1, and Fe Ka1) for each specimen using the function
`penepma_get_intensity_z_s0_s1()` from the
[`rpenepma`](https://github.com/jrminter/rpemepma) package on github.
This function returns the intensity data for the transitions as a
`tibble`. A tibble is an updated dataframe in R that may be processed
using the pipe operator (`%>%`) that is best read as "and then...".


```{r createIntensTibbles}
int_tib_c_std <- penepma_get_intensities(c_std)
int_tib_fe_std_10 <- penepma_get_intensities(fe_std_10)
int_tib_fe_std_p1 <- penepma_get_intensities(fe_std_p1)
int_tib_c200nm_fe_10 <- penepma_get_intensities(c200nm_fe_10)
int_tib_c200nm_fe_p1 <- penepma_get_intensities(c200nm_fe_p1)
int_tib_c20nm_fe_10 <- penepma_get_intensities(c20nm_fe_10)
int_tib_c20nm_fe_p1 <- penepma_get_intensities(c20nm_fe_p1)
```


# Calculate the K-ratios 

Now that we have the intensities, we can compute the k-ratios.
These were computed in the code chunk named `calculateKRs`. The desired
K-ratio was computed by the function `penepma_calc_k_ratio` exported by
the `rpenepma` package. This function computes the desired K-ratio for
the specified transition by extracting the required intensities. The
function also computes the uncertainty for the k-ratio, both as a
value and as a percentage of the k-ratio. To aid in preparing a
summary table, the function returns the sample ID and label for the
transition supplied by the user. See the output below.

```{r calculateKRs}
ret_200_FeK_10 <- penepma_calc_k_ratio("200 nm C on Fe rho=10",
                                  "FeKa",
                                  int_tib_c200nm_fe_10,
                                  int_tib_fe_std_10,
                                  26, 1, 4, n_rnd=4)
ret_200_C_K_10 <- penepma_calc_k_ratio("200 nm C on Fe rho=10", "C Ka",
                                  int_tib_c200nm_fe_10, int_tib_c_std,
                                  6, 1, 4, n_rnd=4)
ret_200_FeL_10 <- penepma_calc_k_ratio("200 nm C on Fe rho=10", "FeLa",
                                  int_tib_c200nm_fe_10, int_tib_fe_std_10,
                                  26, 4, 9, n_rnd=4)

ret_200_FeK_p1 <- penepma_calc_k_ratio("200 nm C on Fe rho=0.1", "FeKa",
                                  int_tib_c200nm_fe_p1, int_tib_fe_std_p1,
                                  26, 1, 4, n_rnd=4)
ret_200_C_K_p1 <- penepma_calc_k_ratio("200 nm C on Fe rho=0.1", "C Ka",
                                  int_tib_c200nm_fe_p1, int_tib_c_std,
                                  6, 1, 4, n_rnd=4)
ret_200_FeL_p1 <- penepma_calc_k_ratio("200 nm C on Fe rho=0.1", "FeLa",
                                  int_tib_c200nm_fe_p1, int_tib_fe_std_p1,
                                  26, 4, 9, n_rnd=4)


ret_20_FeK_10 <- penepma_calc_k_ratio("20 nm C on Fe rho=10",
                                  "FeKa",
                                  int_tib_c20nm_fe_10,
                                  int_tib_fe_std_10,
                                  26, 1, 4, n_rnd=4)
ret_20_C_K_10 <- penepma_calc_k_ratio("20 nm C on Fe rho=10", "C Ka",
                                  int_tib_c20nm_fe_10, int_tib_c_std,
                                  6, 1, 4, n_rnd=4)
ret_20_FeL_10 <- penepma_calc_k_ratio("20 nm C on Fe rho=10", "FeLa",
                                  int_tib_c20nm_fe_10, int_tib_fe_std_10,
                                  26, 4, 9, n_rnd=4)

ret_20_FeK_p1 <- penepma_calc_k_ratio("20 nm C on Fe rho=0.1", "FeKa",
                                  int_tib_c20nm_fe_p1, int_tib_fe_std_p1,
                                  26, 1, 4, n_rnd=4)
ret_20_C_K_p1 <- penepma_calc_k_ratio("20 nm C on Fe rho=0.1", "C Ka",
                                  int_tib_c20nm_fe_p1, int_tib_c_std,
                                  6, 1, 4, n_rnd=4)
ret_20_FeL_p1 <- penepma_calc_k_ratio("20 nm C on Fe rho=0.1", "FeLa",
                                  int_tib_c20nm_fe_p1, int_tib_fe_std_p1,
                                  26, 4, 9, n_rnd=4)


pander(rbind(ret_20_C_K_10, ret_20_FeL_10, ret_20_FeK_10,
             ret_20_C_K_p1, ret_20_FeL_p1, ret_20_FeK_p1,
             ret_200_C_K_10, ret_200_FeL_10, ret_200_FeK_10,
             ret_200_C_K_p1, ret_200_FeL_p1, ret_200_FeK_p1))
```
# Summary

The 20 nm C sample is sufficiently thin that even with an 8 hour
simulation time the FeKa K-ratio for the specimen is greater than 1.
The difficultly is that the relative uncertainties are between 4.9
and 13.6 percent. Perhaps there are changes in simulation parameters
that could permit more trajectories in a shorter time. I am not sure
what biases these changes could introduce.
