---
title: "20-nm-C-250-nm-ZnO-Silica-15kV-reflin-2e-1"
author: "John Minter"
date: 'Started 2018-08-07, Last modified: 2018-08-11'
output:
  html_document:
    css: ../../../../inc/jm.css
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
abstract: |
  The objective was to measure the time required for this simulation using the
  `Zn L3-M4` transition to reach the `REFLIN` criterion of an uncertainty of
  $2.0e^{-1}$.  This was a fairly weak transition (0.1140) compared to the
  stronger `Zn L3-M5` transition (1.0). Still, I want to see how the simulation
  scales with the value of the uncertainty criterion.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
  comment = NA,
  fig.align = "centre",
  fig.height = 4,
  message = FALSE,
  warning = FALSE,
  error = FALSE)
```

# Set up for the analysis

**Note**: This simulation used the same `MSIMPA` parameters for all layers
and a `REFLIN` parameter (`2.E-1`) to finish.

```
RSEED  -10   1                 [Seeds of the random-number generator]
REFLIN 30040800 1 2.E-1     [Zn L3-M4 IZ*1e6+S1*1e4+S2*1e2,dete,tol.]
NSIMSH 1.0e8                    [Desired number of simulated showers]
TIME   1.0e6                       [Allotted simulation time, in sec]
```

First load the libraries we need...

```{r loadPackages}
library(rpenepma)
library(dplyr)
library(pander)
panderOptions('table.split.table', Inf)
library(ggplot2)
```

Next, set the paths and constants that we need...

```{r setPaths}
moved_data  <- TRUE  # set to TRUE when simu is done and .rda with data is moved...
e0          <- 15 # kV
int_lo_lim  <- 1.0e-1
int_hi_lim  <- 1.0e4
reflin_crit <- 2e-1
ori_sim_dir <- "C:/Users/jrminter/Documents/work/penepma16/20-nm-C-250-nm-ZnO-Silica-reflin2"
sim_dir     <- "/Users/jrminter/Desktop/test-plt"
sim_nam     <- "20-nm-C-250-nm-ZnO-Silica-reflin-2e-1"
res_fi      <- "./penepma-res.dat"
int_fi      <- "./pe-intens-01.dat"
spc_fi      <- "./pe-spect-01.dat"
out_ti      <- sprintf("%s-%gkV", sim_nam, e0)
msa_fi      <- sprintf("./%s-%gkV.msa", sim_nam, e0)
```


# Get the current simulation time

```{r calcSimTime}
if (moved_data==FALSE){
  # When we have not moved the sim measure the simulation time.
  # Note copying the files to another directory seems to change the time
  # Test a reload.
  delta_t <- penepma_measure_simulation_time(ori_sim_dir)
  # print(delta_t)
  save(delta_t, file="./delta_t.rda")
  rm(delta_t)
  load("./delta_t.rda")
  msg <- "Analyzed from simulation directory."
} else {
  # When we have moved the simulation, reload the delta_t.rda
  msg <- "Analyzed from data stored in delta_t.rda."
  load("./delta_t.rda")	
}
print(msg)
delta_t_time <- print(delta_t)
```

**Sort summary: this is very short and noisy...**

# Retrieve the number of showers

We want retrieve the number of showers (trajectories) that our `penepma16`
simulation calculated.

```{r calcNumShowers}
options(scipen = -6) # force printing in exponential format
num_showers <- penepma_get_number_of_showers(res_fi)
cat(num_showers)
options(scipen = 3) # reset to default
```

# Compute the total maximum intensity for each element

## Retrieve the data

First retrieve the intensity data as a `tibble` and print a preview.

```{r LoadMaxIntensity}
tib <- penepma_get_intensities(int_fi)
print(tib)
```

What features did we measure?

```{r listFeatures}
print(names(tib))
```

How many transitions did we measure?

```{r calcNumTrans}
print(nrow(tib))
```

Measure our test transition ("Zn L3-M4")

```{r measureZnTrans}
val <- penepma_get_total_intensity_z_transition(tib, 30, "L3", "M4")
print(val)
```

And see how close we are to the `REFLIN` criterion

```{r calcRefCrit}
cur_ref_crit <- val$Int.se/val$Int.mu

rv <- c(cur_ref_crit, 100*reflin_crit/cur_ref_crit)
names(rv) <-c("value", "target %")
print(rv)
```

Let's estimate the remaining time

```{r estimateTimeRemaining}
pct <- rv[2]/100.
names(pct) <- c()
tr <- delta_t*(1.0-pct)
print(tr)
```

Let's store a `.rda` file with `REFLIN` info

```{r storeReflinInfo}

tmp_v <- data.frame(as.character(sim_nam),
                    as.numeric(reflin_crit), as.character.POSIXt(delta_t_time),
                    as.numeric(num_showers), as.numeric(cur_ref_crit),
                    as.numeric(100*reflin_crit/cur_ref_crit))
names(tmp_v) <- c("name", "reflin", "time", "showers",
                  "value", "target %")

tmp_v <- as_tibble(tmp_v)
tmp_v$name <- as.character(tmp_v$name)
C_20_nm_ZnO_250_nm_SiO2_15kV_reflin_2e_m1_ro <- tmp_v
rm(tmp_v)

save(C_20_nm_ZnO_250_nm_SiO2_15kV_reflin_2e_m1_ro,
     file="./C_20_nm_ZnO_250_nm_SiO2_15kV_reflin_2e_m1_ro.rda")
rm(C_20_nm_ZnO_250_nm_SiO2_15kV_reflin_2e_m1_ro)
load("./C_20_nm_ZnO_250_nm_SiO2_15kV_reflin_2e_m1_ro.rda")
pander(C_20_nm_ZnO_250_nm_SiO2_15kV_reflin_2e_m1_ro)
```


## Compute the maximum total intensity for each element

1. Compute the maximum total intensity for each element.
2. Stack each row of data together into a tibble.
3. Prepend the sample ID to the data.
4. Print a well-formatted a table using `pander`.
5. Save the final tibble so it can be r-loaded later

```{r calcNaxTotIntensity}
# 1
C_t   <- penepma_get_max_total_intensity_z(tib, 6)
O_t   <- penepma_get_max_total_intensity_z(tib, 8)
Zn_t  <- penepma_get_max_total_intensity_z(tib, 30)
Si_t  <- penepma_get_max_total_intensity_z(tib, 14)
# 2
tot_int <- dplyr::bind_rows(C_t, O_t, Zn_t, Si_t)
#3 
C_20_nm_ZnO_250_nm_SiO2_15kV_reflin_2e_m1 <- prepend_sample_id_max_int_tib(tot_int, out_ti)
# 4
pander(C_20_nm_ZnO_250_nm_SiO2_15kV_reflin_2e_m1)
# 5
save(C_20_nm_ZnO_250_nm_SiO2_15kV_reflin_2e_m1,
     file="C_20_nm_ZnO_250_nm_SiO2_15kV_reflin_2e_m1.rda")
# delete and reload be sure it worked...
rm(C_20_nm_ZnO_250_nm_SiO2_15kV_refli2n)
load("./C_20_nm_ZnO_250_nm_SiO2_15kV_reflin_2e_m1.rda")
```

# Process the spectra data

## Load the raw data

```{r loadSpectraData}
tib <- penepma_read_raw_data(spc_fi,min_intensity_clip=5.0e-10)
rownames(df) <- c()
```

Examine the start:

```{r examineHead}
pander(head(tib))
```

Examine the end:

```{r examineTail}
pander(tail(tib))
```
## Plot the spectrum.

There is a very large dynamic
range for both the **probability density** and the **uncertainty**.
Penepma sets a lower limit for data at **1.0e-35**. Missing values are
set to zero. We want to remove values from the dataframe that are
below a useful limit. We do this below and plot a copy of the 
dataframe that is limited to the useful values.


```{r plotLogIntensSpectrum, fig.width=8, warning=FALSE}
m_t <- max(pretty(tib$mu))

plt <- ggplot(tib, aes(x = keV, y = mu)) +
       geom_line() + 
       scale_x_continuous(breaks = seq(from = 0, to = e0-5, by = 1),
                          limits = c(0,e0-5)) +
       scale_y_log10(limits = c(int_lo_lim, m_t)) +
       xlab(label="X-ray energy [keV]") +
       ylab(label="log(probability density)") +
       # (1/(eV*sr*electron)") +
       ggtitle(out_ti) +
       theme(axis.text=element_text(size=12),
             axis.title=element_text(size=12),
             # center the title
             plot.title = element_text(hjust = 0.5))
       
print(plt)
```
And on a linear intensity scale...

```{r plotLinIntensSpectrum, fig.width=8, warning=FALSE}

plt <- ggplot(tib, aes(x = keV, y = mu)) +
       geom_line() + 
       scale_x_continuous(breaks = seq(from = 0, to = e0-5, by = 1),
                          limits = c(0,e0-5)) +
       scale_y_continuous(limits = c(int_lo_lim, m_t)) +
       xlab(label="X-ray energy [keV]") +
       ylab(label="Probability density") +
       # (1/(eV*sr*electron)") +
       ggtitle(out_ti) +
       theme(axis.text=element_text(size=12),
             axis.title=element_text(size=12),
             # center the title
             plot.title = element_text(hjust = 0.5))
       
print(plt)
```

## Write a spectrum file in MSA format

```{r writeMsa}
penepma_to_msa(spc_fi, msa_fi,e0, out_ti)
```


