---
title: "20-nm-C-250-nm-ZnO-Silica-15kV"
author: "John Minter"
date: "2018-08-06"
output:
  html_document:
    css: ../../../inc/jm.css
    toc: true
    number_sections: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
  comment = NA,
  fig.align = "centre",
  fig.height = 4,
  message = FALSE,
  warning = FALSE,
  error = FALSE)
```

# Set up for the analysis

**Note**: This version used the same `MSIMPA` parameters for all layers.

First load the libraries we need...

```{r loadPackages}
library(rpenepma)
library(dplyr)
library(pander)
library(ggplot2)
```


Next, set the paths and constants that we need... **Note**: We expect
to run this script from a directory within the `monte-carlo-simulations`
git project folder (cloned from
[here](https://github.com/jrminter/monte-carlo-simulations)) and expect that
to be in the user's `git` directory, the path of which is stored in the
`GIT_HOME` environment variable.

```{r setPaths}
e0         <- 15 # kV
int_lo_lim <- 1.0e-1
int_hi_lim <- 1.0e4
git_home   <- Sys.getenv('GIT_HOME')
rel_dir    <- "monte-carlo-simulations/penepma/penepma16"
sim_nam    <- "20-nm-C-250-nm-ZnO-Silica"
sim_dir    <- sprintf("%s/%s/%s-%gkV", git_home, rel_dir, sim_nam, e0)

res_fi  <- "./penepma-res.dat"
int_fi  <- "./pe-intens-01.dat"
spc_fi  <- "./pe-spect-01.dat"
out_ti  <- sprintf("%s-%gkV", sim_nam, e0)
msa_fi  <- sprintf("./%s-%gkV.msa", sim_nam, e0)
# print(getwd())
# wd <- sprintf("%s/%s", sim_dir, sim_nam)

```

# Retrieve the number of showers

We want retrieve the number of showers (trajectories) that our `penepma16`
simulation calculated.

```{r calcNumShowers}
options(scipen = -6) # force printing in exponetial format
num_showers <- penepma_get_number_of_showers(res_fi)
cat(num_showers)
options(scipen = 3) # reset to default
```

# Compute the total maximum intensity for each element

## Retrieve the data

First retrieve the intensity data as a `tibble` and print a preview.

```{r LoadMaxIntensity}
tib <- penepma_get_intensities(int_fi)
print(tib)
```

What features did we measure?

```{r listFeatures}
print(names(tib))
```

How many transitions did we measure?

```{r calcNumTrans}
print(nrow(tib))
```

## Compute the maximum total intensity for each element

1. Compute the maximum total intensity for each element.
2. Stack each row of data together into a tibble.
3. Prepend the sample ID to the data.
4. Print a well-formatted a table using `pander`.
5. Save the final tibble so it can be r-loaded later

```{r calcNaxTotIntensity}
# 1
C_t   <- penepma_get_max_total_intensity_z(tib, 6)
O_t   <- penepma_get_max_total_intensity_z(tib, 8)
Zn_t  <- penepma_get_max_total_intensity_z(tib, 30)
Si_t  <- penepma_get_max_total_intensity_z(tib, 14)
# 2
tot_int <- dplyr::bind_rows(C_t, O_t, Zn_t, Si_t)
#3 
C_20_nm_ZnO_250_nm_SiO2_15kV <- prepend_sample_id_max_int_tib(tot_int, out_ti)
# 4
panderOptions('table.split.table', Inf)
pander(C_20_nm_ZnO_250_nm_SiO2_15kV)
# 5
save(C_20_nm_ZnO_250_nm_SiO2_15kV, file="C_20_nm_ZnO_250_nm_SiO2_15kV.rda")
# delete and reload be sure it worked...
rm(C_20_nm_ZnO_250_nm_SiO2_15kV)
load("./C_20_nm_ZnO_250_nm_SiO2_15kV.rda")
```

# Process the spectra data

## Load the raw data

```{r loadSpectraData}
tib <- penepma_read_raw_data(spc_fi)
rownames(df) <- c()
```

Examine the start:

```{r examineHead}
pander(head(tib))
```

Examine the end:

```{r examineTail}
pander(tail(tib))
```
## Plot the spectrum.

There is a very large dynamic
range for both the **probability density** and the **uncertainty**.
Penepma sets a lower limit for data at **1.0e-35**. Missing values are
set to zero. We want to remove values from the dataframe that are
below a useful limit. We do this below and plot a copy of the 
dataframe that is limited to the useful values.


```{r plotLogIntensSpectrum, fig.width=8, warning=FALSE}
plt <- ggplot(tib, aes(x = keV, y = pd.mu)) +
       geom_line() + 
       scale_x_continuous(breaks = seq(from = 0, to = e0-5, by = 1),
                          limits = c(0,e0-5)) +
       scale_y_log10() +
       xlab(label="X-ray energy [keV]") +
       ylab(label="log(probability density)") +
       # (1/(eV*sr*electron)") +
       ggtitle(out_ti) +
       theme(axis.text=element_text(size=12),
             axis.title=element_text(size=12),
             # center the title
             plot.title = element_text(hjust = 0.5))
       
print(plt)
```
And on a linear intensity scale...

```{r plotLinIntensSpectrum, fig.width=8, warning=FALSE}
plt <- ggplot(tib, aes(x = keV, y = pd.mu)) +
       geom_line() + 
       scale_x_continuous(breaks = seq(from = 0, to = e0-5, by = 1),
                          limits = c(0,e0-5)) +
       scale_y_continuous() +
       xlab(label="X-ray energy [keV]") +
       ylab(label="Probability density") +
       # (1/(eV*sr*electron)") +
       ggtitle(out_ti) +
       theme(axis.text=element_text(size=12),
             axis.title=element_text(size=12),
             # center the title
             plot.title = element_text(hjust = 0.5))
       
print(plt)
```

## Write a spectrum file in MSA format

```{r writeMsa}
penepma_to_msa(spc_fi, msa_fi,e0, out_ti)
```


